---
title: "Final Project"
author: "Holly Finertie"
date: "5/4/2020"
output:
  html_document: default
---

## COVID-19 in California by County and County Characterisitics


#### Step1: Load Packages
```{r}
library(tidyverse)
require(sf)
require(rgdal)
require(spdep)
require(maptools)
library(maps)
library(leaflet)
```

#### Step 2: Import CA COVID-19 Cases by county

Data on cumulative cases and deaths from COVID-19 by county was downloaded from [NYTimes Github](https://github.com/nytimes/covid-19-data). The results were filtered to include total cases and deaths in California counties as of 5/1/2020. Data were missing from 4 counties: Lassen, Modoc, Sierra, and Trinity. 

```{r}
covid_nyt = read_csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv") %>% 
  filter(date == "2020-05-01",
         state == "California") %>% 
  select(-fips, -state, -date) %>% 
  mutate(
    county = as.character(county)
  )
```


#### Step 3: Import CA Demographic, Household, and Commuting Data

Data on demographics, average household size, and commuter characteristics by California county were downloaded from [Census Data](https://data.census.gov/cedsci/). This data is from the 2013-2017 American Community Survey (ACS) 5-year estimates. 

```{r}
demo_acs = read_csv("./california_county_demo.csv")
household_acs = read_csv("./california_county_household.csv")
commute_acs = read_csv("./california_county_commute.csv")

data_acs = merge(demo_acs, household_acs, by = c("id", "county")) %>% 
  merge(commute_acs, by = c("id", "county")) %>% 
  janitor::clean_names() %>% 
  mutate(
    county = as.character(strsplit(county, " County, California"))
  ) %>% 
  select(-id)

```

#### Step 4: Merge Data Files and create COVID per 100k stats

Create final non-spatial data frame with all CA counties including COVID-19 and ACS data. Created variables to represent cases and deaths per 100,000 people. 

```{r}
covid_final_df = left_join(data_acs, covid_nyt, by = "county") %>% 
  mutate(
    cases_per_100k = round((cases/population)*100000, digits = 2), 
    deaths_per_100k = round((deaths/population)*100000, digits = 2)
  ) %>% 
  mutate(
    county = str_to_lower(county))

row.names(covid_final_df) = covid_final_df$county
```

#### Step 5: Import CA Counties map

Imported California map with county polygons from maps package. 

```{r}
ca_county = map('county', 'california', fill = TRUE, col = palette())

county_names = strsplit(ca_county$names,",")

map_id = as.character(tolower(sapply(county_names, function(x) x[2])))

ca_map = map2SpatialPolygons(ca_county, IDs = map_id, 
                                proj4string = CRS("+init=epsg:3310"))
plot(ca_map)
```


#### Step 6: Create Shape File with all Data

Created shapefile that merged non-spatial dataframe with CA spatial polygon file. Wrote and saved shapefile. 

```{r}
ca_spatial_df = SpatialPolygonsDataFrame(ca_map, covid_final_df)

summary(ca_spatial_df)

writeOGR(ca_spatial_df, 
         dsn = "data", 
         layer = "CA_COVID",
         overwrite_layer = TRUE,
         driver = "ESRI Shapefile")

ca_covid = readOGR(dsn = "data",
                  layer = "CA_COVID") 

```


##### Step 7: Create Leaflet

Using Leaflet package, created an interactive map with COVID-19 information and county characteristics. 

```{r}
binpal = colorBin("Reds", domain = ca_covid$cs__100, 5)

ca_covid$county = str_to_title(ca_covid$county)

names(ca_covid)

ca_map = leaflet(ca_covid) %>% 
  setView(-119.417,36.778, 5) %>%
  addTiles() %>% 
  addPolygons(fillColor = ~binpal(cs__100),
                             weight = 2,
                             color = 'black',
                             fillOpacity = 0.5,
                             popup = paste("<u>", ca_covid$county, "County</u><br>", 
                                           "COVID-19 Cases per 100k (5/1/2020):", ca_covid$cs__100, "<br>",
                                           "COVID-19 Deaths per 100k (5/1/2020):", ca_covid$dt__100, "<br>",
                                           "Percent 65+:", ca_covid$prcn_65, "<br>",
                                           "Percent White:", ca_covid$prcnt_w, "<br>",
                                           "Average Household Size:", ca_covid$hshld_s, "<br>", 
                                           "Percent Commute Public Transportation", ca_covid$prcnt_p_, "<br>"
                                        ))

ca_map

require(htmlwidgets)
saveWidget(ca_map, file = "CA_COVID-19.html")
```



