---
title: "Final Project"
author: "Holly Finertie"
date: "5/4/2020"
output: html_document
---

#### Step1: Import packages
```{r}
library(tidyverse)
require(sf)
require(rgdal)
require(spdep)
require(maptools)
```

#### Step 2: Import CA COVID-19 Cases by county

Data downloaded from [NYTimes Github](https://github.com/nytimes/covid-19-data). The NYTimes had access to 54 counties in California and filtered the results to cases as of May 1, 2020. Cases are crude numbers. 
```{r}
covid = read_csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv") %>% 
  filter(date == "2020-05-01",
         state == "California")

```


#### Step 3: Import CA Demographic, Household, and Commuting Data

Data downloaded from [Census Data](https://data.census.gov/cedsci/) on demographics (age, sex),average household size, and percent of people who drive or use public transportation to get to work. 


```{r}
demo = read_csv("./california_county_demo.csv")
household = read_csv("./california_county_household.csv")
commute = read_csv("./california_county_commute.csv")

non_covid_data = merge(demo, household, by = c("id", "county")) %>% 
  merge(commute, by = c("id", "county")) %>% 
  janitor::clean_names()

```

